---
title: "Project 1 RR"
author: "Shadrack Badia"
date: "4/26/2021"
output: html_document
---


```{r}
# create a temporary file on the local disk
tempf <- tempfile()
# download data to the temporary file
download.file( "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" , tempf , mode = 'wb' )

# unzip the contents 
data <- unzip( tempf , exdir = tempdir() )

data1 <- read.csv(data, sep = ",",header = TRUE)


```




```{r}
library(ggplot2)
library(dplyr)
library(lattice)
```


```{r}
#Looking at the strings and how data is composed
str(data1)
```



```{r}
#Looking at a summary of Numerical data Variables
summary(data1)
```





####QUESTION 1


###1.Calculate the total number of steps taken per day


```{r}
stepsperday<-aggregate(steps~date, data1, sum)
```


### 2.f you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

```{r}
hist(stepsperday$steps, xlab="Group by Total Number of Steps per day", ylab="Number of Days", main="Total Number of Steps taken each day")
```


The MEAN

```{r}
meanstp<-mean(stepsperday$steps)
meanstp
```

3.The Median

```{r}
medianStp<-median(stepsperday$steps)
medianStp
```


4.Time series plot of the average number of steps taken

```{r}
#the average daily activity pattern?
avgdaily<-aggregate(steps ~ interval, data1, mean)
```


```{r}
with(avgdaily, plot(interval, steps, type = "l"))
```

5.The 5-minute interval that, on average, contains the maximum number of steps
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?



```{r}
avgdaily[which.max(avgdaily[,2]),1]
```

6.Code to describe and show a strategy for imputing missing data

Calculate Missing Values
```{r}
#Misssing data
vis_miss(data1)
```


```{r}
gg_miss_var(data1) + labs(y = "Look at all the missing ones")
```

```{r}
#total number of missing variables are all 0 values
n_distinct(data1$steps)
```


```{r}
miss_case_table(data1)
```

13.1% representing 2304 consist of all missing valiies of the dataset

###Filling missing values

Filling the values using the mean of the steps mean is x

```{r}
x <- mean(stepsperday$steps)/365
x
```
```{r}
data2<-data1
data2[missingIndex,1]<-x
head(data2)
```

Finding the total number of steps each day after missing values are imputed and making histogram:

```{r}
newstepsperday<-aggregate(steps~date, data2, sum)
hist(newstepsperday$steps, xlab="Group by Total Number of Steps per day", ylab="Number of Days", main="Number of Steps taken each day after missing values are imputed")
```

```{r}
newmean <- mean(newstepsperday$steps)
newmean
```

```{r}
newmedian <- median(newstepsperday$steps)
newmedian
```
### Are there differences in activity patterns between weekdays and weekends?
For this part the \color{red}{\verb|weekdays()|}weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
data2$date<-as.Date(data2$date)
```


8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
data3<-data2%>%
        mutate(dayType= ifelse(weekdays(data2$date)=="Saturday" | weekdays(data2$date)=="Sunday", "Weekend", "Weekday"))
head(data3)
```


Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
#Let z represent the Average Step By  Day Type And Interval
z<-data3 %>%
  group_by(dayType, interval) %>%
  summarize(averageStepByDay=sum(steps))

head(z)
```

```{r}
with(z, 
      xyplot(averageStepByDay ~ interval | dayType, 
      type = "l",      
      main = "The Total Number of Steps within Intervals by dayType",
      xlab = "Daily Intervals",
      ylab = "Average Number of Steps"))
```




